---
title: "STAT 4730 Lab 2"
author: "Ben Lehmann"
output: pdf_document
---

# Lab Assignment 2

Please make sure to insert a code chunk (Ctrl + Alt + I) and write your code in it, **under each question**, if any coding is needed. 

In this assignment we are going to simulate two data sets, merge them together, and conduct a simple analysis.


```{r load libraries}

library(dplyr)

```


## Question 1

1. Set the seed using `set.seed(2025)`.

```{r}

# Write R code here

# Remember to compile your code use the "knit" button in RStudio or use
# the rmarkdown::render() function.
set.seed(2025)

```

## Question 2

2. Create a data frame called `data_1` with the following columns:
   * `id` - integers that go from 1 to 10
   * `num1` - 10 numbers that come from a normal distribution with a mean of 0
     and a standard deviation of 1
   * `num2` - 10 numbers that come from a normal distribution with a mean of 2
     and a standard deviation of 3
   * `num3` - 10 numbers that come from a normal distribution with a mean of 1
     and a standard deviation of 2.

```{r}
data_1 <- data.frame(
  id = 1:10,
  num1 = rnorm(10, mean = 0, sd = 1),
  num2 = rnorm(10, mean = 2, sd = 3),
  num3 = rnorm(10, mean = 1, sd = 2)
)
data_1
```



## Question 3

3. Create a data frame called `data_2` with the following columns:
   * `id` - integers that go from 1 to 10
   * `num4` - 10 numbers that come from a uniform distribution with a minimum
     of -1 and a max of 1. See the help for `runif`.
   * `num5` - 10 numbers that come from a uniform distribution with a minimum
     of 0 and a max of 3.
     

```{r}
data_2 <- data.frame(
  id = 1:10,
  num4 = runif(10, min = -1, max = 1),
  num5 = runif(10, min = 0, max = 3)
)
data_2
```


## Question 4

4. Join `data_1` and `data_2` and keep all available columns. **Display** the new data.

```{r}
merged_data <- inner_join(data_1, data_2, by = "id")  #Use inner_join
merged_data
```



## Question 5

5. Find the mean of each of the columns `num1` through `num5` from the data set
   in Question 4. Do NOT report the mean of the `id` column. (There are multiple
   ways to correctly solve this problem.)
   
```{r}
column_means <- colMeans(merged_data[, c("num1", "num2", "num3", "num4", "num5")])
column_means
```


# Further Exploration (Required for 573 Students only)

1. Last week, we worked with data that had been grouped. Can we sample from 
   grouped data? See if you can sample 1 row from the `starwars` data 
   set of each unique species.


2. If you look at the help function for the built-in `mean` function, it has the
   option of ignoring `NA` values. How could you update our `mean_fun` so that
   it can ignore `NA` values? **Test** the result by applying the function to any vector that includes `NA` values.

  