---
title: "STAT 4730 Lab 1"
author: "Ben Lehmann"
output: pdf_document
---

The following document shows a good way to answer the homework questions.

# Lab 1

Before we answer any question it is good practice to load the required libraries.

```{r load libs}

library(dplyr)
library(readr)
library(tidyr)

```

## Question 1

1. Load the `agpop.dat` data set. (It is separated by spaces. Make sure to
   keep the first row of data points not as column names.) Hint: Look at the help file for `readr::read_delim`.

   Write your answer here

```{r block title}

# Write R code here

# Remember to compile your code use the "knit" button in RStudio or use
# the rmarkdown::render() function.

agpop <- read_delim("agpop.dat",delim = " ", col_names = FALSE, trim_ws = TRUE)

```

## Question 2

2. Create a character vector with the following names (in this order):
   county, state, acres92, acres87, acres82, farms92, farms87, farms82,
   largef92, largef87, largef82, smallf92, smallf87, smallf82, and region.
   
```{r}
column_names <- c(
  "county", "state", "acres92", "acres87", "acres82", 
  "farms92", "farms87", "farms82", 
  "largef92", "largef87", "largef82", 
  "smallf92", "smallf87", "smallf82", "region"
)
```


## Question 3

3. Change the names of the `agpop` data set to the names in Question 2. (`names` function)

```{r}
names(agpop) <- column_names
agpop
```


## Question 4

4. Keep the following columns in the `agpop` data: county, state, acres92,
   acres87, and region. Assign this new data frame to another name. (`select` function)


```{r}
agpop_selected <- agpop %>% select(county, state, acres92, acres87, region)
agpop_selected
```


## Question 5

5. Get the number of observations in the `agpop` data set.

```{r}
num_observations <- nrow(agpop)
num_observations
```

```{r}
observations_new <- nrow(agpop_selected)
observations_new
```


## Question 6

6. Remove the observations with the `acres92` less than `acres87` and assign this new data set another name. (`filter` function)

We can filter from the dataset from Question 4 (With only 5 variables)
```{r}
agpop_filtered <- agpop_selected %>% 
  filter(acres92 < acres87)

agpop_filtered
```

We can also filter from the original Dataset with all 15 variables
```{r}
agpop_filter <- agpop %>% 
  filter(acres92 < acres87)

agpop_filter
```

Filtered between the two different datasets, still got the same number of rows

## Question 7

7. Get the number of observations in this new smaller data set.

```{r}
num_filtered_observations <- nrow(agpop_filtered)
num_filtered_observations
```


## Question 8

8. What do your results tell you about farm sizes between 1987 and 1992?

We can say that after filtering our data we can see that some counties saw a decrease in acres sizes farmed between 1987 and 1992

# Further Exploration Solution (Required for 573 Students Only)

The `agpop.dat` file is a SAS data set. In SAS, the number "-99" is used to
indicate a missing (`NA`) value. Find the number of observations in the `agpop`
data set with all of the `NA` values removed.





